---
title: "proj1.Rmd"
output: 
        html_document:
                toc: true
                toc_float: true
                number_sections: true
                theme: darkly
---
# About {-}

This document includes all the notes, codes, and thought processes for week 1 project

## Submission Notes

repo link: https://github.com/tanyayt/coursera_public/tree/master/exploratory_data_analysis_w1_proj



## Plan

        * 1. Read and understand the data discription
        * 2. Use this file as the ultimate analysis log
        * 3. Export using purl, to R code for submission
        
## Data Source

[ref](https://www.coursera.org/learn/exploratory-data-analysis/peer/ylVFo/course-project-1)

*Description*: Measurements of electric power consumption in one household with a one-minute sampling rate over a period of almost 4 years. Different electrical quantities and some sub-metering values are available.

*Variables*:

        - Date: Date in format dd/mm/yyyy
        - Time: time in format hh:mm:ss
        - Global_active_power: household global min-avg active power (kw)
        - Global_reactive_power: household global min-avg reactive power (kw)
        - Voltage: minute-averaged voltage (in V)
        - Global_intensity: household global min-avg current intensity (Amp)
        - Sub_metering_1: energy sub-metering No. 1 (in watt-hour of active energy). It corresponds to the kitchen, containing mainly a dishwasher, an oven and a microwave (hot plates are not electric but gas powered).
        - Sub_metering_2: energy sub-metering No. 2 (in watt-hour of active energy). It corresponds to the laundry room, containing a washing-machine, a tumble-drier, a refrigerator and a light.
        - Sub_metering_3: energy sub-metering No. 3 (in watt-hour of active energy). It corresponds to an electric water-heater and an air-conditioner.
        
# Download and Import Data
## Download

change eval to TRUE, if you haven't downloaded the file
```{r download-data,eval=FALSE}
library(downloader)
downloader::download("https://d396qusza40orc.cloudfront.net/exdata%2Fdata%2Fhousehold_power_consumption.zip",destfile="proj1_power_consumption.zip")
unzip("proj1_power_consumption.zip")
```

## Import and Clean 

### sample_data
```{r import_sample}
sample_data<-data.frame()
sample_data<-read.table("household_power_consumption.txt",skip=grep("1/2/2007",readLines("household_power_consumption.txt")),nrows = 100,sep = ";",header=TRUE)
head(sample_data)
```

from the output, we know the date for Feb 1, 2017 is actually coded as "1/2/2007"

```{r import_sample_2}
sample_data<-data.frame()
sample_data<-read.table("household_power_consumption.txt",skip=grep("2/2/2007",readLines("household_power_consumption.txt")),nrows = 100,sep = ";",header=TRUE,na.strings = c("?",""))
head(sample_data)
```
as expected, Feb 2, 2017 is coded as "2/2/2007"

### selectlively import data in consump dataframe
```{r select_import_consump}
# prepration - uncomment if not installed
# install.packages("sqldf")
library(sqldf)
consump<-data.frame()
consump<-read.csv.sql(
        file="household_power_consumption.txt",
        sql = "SELECT * FROM file WHERE Date IN ('1/2/2007','2/2/2007') ",
        header=TRUE,
        sep=";",
        row.names=FALSE,
        colClasses="character"
        )
head(consump)

```

```{r examine_data}
# must use fixed = TRUE, since "?" is a special character 
# or use "\\?""
length(grep("?",consump,fixed = TRUE)) 

```

```{r backup}
write.csv(consump,file="consump_data_20070201_20070202.csv",row.names = FALSE) 
#create a backup copy
```

```{r backup_reassign}
consump<-read.csv("consump_data_20070201_20070202.csv",header = TRUE,sep=",")

```

```{r clean_consump}
library(dplyr)
consump$Date<-as.Date(consump$Date,"%d/%m/%Y") #must use capital Y for                                                      #four-digit year
consump<-dplyr::mutate(consump,date_time=paste0(Date," ",Time))
consump$date_time<-strptime(consump$date_time,"%Y-%m-%d %H:%M:%S")
summary(consump$date_time)
head(consump$date_time)
```

up to this point, the data is clean

```{r examine_columns}
summary(consump)

```



# Plot

The overall goal is to examine how household energy useage varies over the two days. 

## Plot 1
```{r plot1}
png("plot1.png",width =480, height = 480,units = "px")
plot1<-function(){
        hist(consump$Global_active_power,main = "Global Active Power",
     col = "red",xlab = "Global Active Power (killowatts)")
}
plot1()
dev.off()
```

## Plot 2
```{r plot2}
png("plot2.png",width =480, height = 480,units = "px")
plot2<-function(){plot(consump$date_time,consump$Global_active_power,type="l",xlab ="",ylab = "Global Active Power (kilowatts" )}
plot2()

dev.off()

```

## Plot 3

```{r plot3}
png("plot3.png",width =480, height = 480,units = "px")
plot3<-function(){
        plot(consump$date_time,consump$Sub_metering_1,
             type = "l",
             col="black",
             xlab="",
             ylab="Energy sub metering")
        lines(consump$date_time,consump$Sub_metering_2,col="red")
        lines(consump$date_time,consump$Sub_metering_3,col="blue")
                legend("topright",
       legend = c("Sub_metering_1","Sub_metering_2","Sub_metering_3"),
       col=c("black","red","blue"),
       lty = 1
        )
}
plot3
dev.off()
```

## Plot 4
```{r plot4}
png("plot4.png",width =480, height = 480,units = "px")

plot5<-function(){
        plot(consump$date_time,consump$Voltage,type = "l",
        xlab="datetime",
        ylab="Voltage")
                }

plot6<-function(){
        plot(consump$date_time,consump$Global_reactive_power,
             type="l",
             xlab="datetime",
             ylab="Global_reactive_power"
             )
}

par(mfrow=c(2,2))
plot2()
plot5()
plot3()
plot6()

dev.off()
```

