---
title: "w2qz.Rmd"
output: 
        html_document:
                toc: true
                toc_float: true
                number_sections: false
                theme: darkly
---
# Question 1 Under the lattice graphics system, what do the primary plotting functions like xyplot() and bwplot() return?

Ans: an object of class "trellis"

# Question 2 What's produced by code

```{r}
library(nlme)
library(lattice)
xyplot(weight ~ Time | Diet, BodyWeight)

```

ans: A set of 3 panels showing the relationship between weight and time for each diet.

# Question 3, Which of the following functions can be used to annotate the panels in a multi-panel lattice plot?

ans: lpoints()

# Question 4， Which of the following is an explanation for why no plot appears?

```{r}
library(lattice)
library(datasets)
data(airquality)
xyplot(Ozone ~ Wind | factor(Month), data = airquality)
```
Ans: The object 'p' has not yet been printed with the appropriate print method.

# Question 5 In the lattice system, which of the following functions can be used to finely control the appearance of all lattice plots?

ans:trellis.par.set()

# Q6 What is ggplot2 an implementation of?
ans： the Grammar of Graphics developed by Leland Wilkinson

# Q7 Load the `airquality' dataset form the datasets package in R
ans: 

```{r}
library(datasets)
data(airquality)
library(ggplot2)
qplot(Wind, Ozone, data = airquality, facets = . ~ factor(Month))
```

ans: qplot(Wind, Ozone, data = airquality, facets = . ~ factor(Month))

# Q8 What is a geom in the ggplot2 system?
ans：a plotting object like point, line, or other shape

# Q9  was expecting a scatterplot of 'votes' and 'rating' to appear. What's the problem?
ans: ggplot does not yet know what type of layer to add to the plot.

#Q10 The following code creates a scatterplot of 'votes' and 'rating' from the movies dataset in the ggplot2 package. After loading the ggplot2 package with the library() function, I can run

ans: qplot(votes, rating, data = movies) + geom_smooth()
```{r}
qplot(Ozone, Wind, data = airquality) + geom_smooth()
```

