---
title: "pa2_lexical scoping"
output: github_document
---
# About
Course: R Programming
Platform: Coursera
Assignment: Programming Assignment 2: Lexical Scoping

# Goal 
Wrte an R function to cache potentially time-consuming computations. 

# Knowledge 

        * <<- operator to assign a value to an object in an environment that's differnt from the current environment 
        
# makeCacheMatrix

This function creates a special matrix object that can cache its inverse

```{r makeCacheMatrix}
makeCacheMatrix <-function (x=matrix()) {
        #set initial value x to an empty matrix
        m <-NULL 
        set <- function (y){
                x <<-y
                m <<-NULL
        }
        
        get <- function ()x # assign value x to get function
        setinverse <-function(solve) {m<<-solve} # assign solve to m 
        getinverse <-function() m
        
        list (setvalue = set, getvalue= get,
              setinverse_value = setinverse,
              getinverse_value = getinverse)
}

```

#cacheSolve

This function calculates the inverse of the matrix x; if the inverse is in the cache, the calculation is not executed rather just give the value in cache 

```{r cacheSolve}
cacheSolve <-function(x){
        m <-x$getinverse()
        if(!is.null(m)){
                message("cached data already existed")
                m
        }
        
        data<-x$get()
        m<- solve(data)
        x$setinverse(m)
        m

}

```


