---
title: "qz2.Rmd"
output: 
        html_document:
                toc: true
                toc_float: true
                number_sections: false
                theme: darkly
---
# Q1

```{r}
x <- c(0.61, 0.93, 0.83, 0.35, 0.54, 0.16, 0.91, 0.62, 0.62)
y <- c(0.67, 0.84, 0.6, 0.18, 0.85, 0.47, 1.1, 0.65, 0.36)
summary(lm(y~x))$coefficient[2,4]
```

note by default, lm returns the two-sided p values; the [2,4] index returns the second coefficient, beta1, and the fourth column value (which is p)

#Q2
the question asks for residul vairation
```{r}
fit<-lm(y~x)
summary(fit)$sigma
```

alternatively, it is also 
```{r}
sqrt((sum(resid(fit)^2)/(length(x)-2)))
```

#Q3 

```{r}
library(datasets)
data("mtcars")
y<-mtcars$mpg
x<-mtcars$wt
fit<-lm(y~x)
meanx<-mean(x)
predict(fit,newdata = data.frame(x=mean(x)),interval = "confidence")
```

#Q4 
```{r}
?mtcars
#weight is per 1000lb 
#ans the estimated expected change in mpg per 1k lb increase in weight 
```


#Q5
```{r}
predict(fit,newdata=data.frame(x=3),interval="prediction") #note this should be prediction interval 
```
ans: 27.57

#Q6 

```{r}
sumcoef<-summary(fit)$coefficients
(sumcoef[2,1]+c(-1,1)*qt(0.975,df=fit$df)*sumcoef[2,2])*2
```


#Q7 

it will be divided by 100 


#Q8
ans: the new intercept would be beta_zero_hat - beta1_hat*c

#Q9

the question asks about the residual variations of two models 
residual variation (the sum of squarefit/n-2) can be obtained by $sigma
```{r}
fit1<-lm(y~x)
fit2<-lm(y~1) # which force the slope to be zero 

summary(fit1)$sigma/summary(fit2)$sigma

```


#Q10 

the residuals must always sum to zero 