---
title: "Quiz 3 - Practical Machine Learning"
author: "Tanya"
date: "10/30/2019"
output: html_document
---
# Package versions
```{r}
packageVersion("AppliedPredictiveModeling")
```
```{r}
packageVersion("caret")
```
```{r}
packageVersion("ElemStatLearn")
```

```{r}
packageVersion("pgmm")
```

```{r}
packageVersion("rpart")
```

```{r load_data}
library(AppliedPredictiveModeling)
data(segmentationOriginal)
```

# Task 1 Subset the data into training and testing
```{r}
table(segmentationOriginal$Case)
```
```{r}
training<-segmentationOriginal[segmentationOriginal$Case=="Train",]
testing<-segmentationOriginal[segmentationOriginal$Case=="Test",]
```

# Task 2 set seed and fit CART to predict Class with rpart method
```{r}
summary(training$Class)
```
```{r}
set.seed(125)
MODEL1<-caret::train(Class~.,data = training,method = "rpart") 
fancyRpartPlot(MODEL1$finalModel)
```

rpart generates classification and regression trees 

for a: PS
b. WS
c. PS
d. can't predict 

#Q2
if K is small, fewer groups in the data set; smaller K, smaller variance less bias; 

ans: The bias is larger and the variance is smaller. Under leave one out cross validation K is equal to the sample size.
 

#Q3 

```{r}
library(pgmm);data(olive)
olive<-olive[,-1]
summary(olive$Area)
```
```{r}
#str(olive)
MODELOLIVE<-train(Area~.,data = olive,method="rpart")
fancyRpartPlot(MODELOLIVE$finalModel)
```

```{r}
NEWDATA<-as.data.frame(t(colMeans(olive)))
predict(MODELOLIVE$finalModel,NEWDATA)
```


#Q4 

prep:
```{r}
library(ElemStatLearn)
data(SAheart)
set.seed(8484)
train = sample(1:dim(SAheart)[1],size=dim(SAheart)[1]/2,replace=F)
trainSA = SAheart[train,]
testSA = SAheart[-train,]
```

build the model
```{r}
trainSA$chd<-as.factor(trainSA$chd)
testSA$chd<-as.factor(testSA$chd)
set.seed(13234)
MODELSA<-train(chd~age+alcohol+obesity+tobacco+typea+ldl,data=trainSA,method="glm",family="binomial")

missClass = function(values,prediction){sum(((prediction > 0.5)*1) != values)/length(values)}

```

misclassification error for training set
```{r}
predictedvalues<-predict(MODELSA$finalModel,trainSA)
missClass(trainSA$chd,predictedvalues)
```

```{r}
predictedvaluesTEST<-predict(MODELSA$finalModel,testSA)
missClass(testSA$chd,predictedvaluesTEST)
```
the closest is training 0.31 amd test 0.35 

#Q5 

```{r}
library(ElemStatLearn)
data(vowel.train)
data(vowel.test)
```

```{r}
vowel.train$y<-as.factor(vowel.train$y)
vowel.test$y<-as.factor(vowel.test$y)
```

```{r}
set.seed(33833)
modelRF <- randomForest(y~.,data=vowel.train)
order(caret::varImp(modelRF),decreasing=TRUE)
```


```{r}
varImpPlot(modelRF)
```


