---
title: "w4_project"
author: "Tanya Tan"
date: "28/08/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About
This file documents the process working through week 4 course project in data science/getting andcleaning data course. 

* Naming Strategy: w4_proj_XXX where XXX stands for the file name

# Load the data and store in w4_project_data.zip

The following code is not in a code chunk on purpose so that it won't download the file in every refresh
```{r download file}
library(downloader)
download("https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip",dest="w4_project_data.zip",mode="wb") #wb stands for write and binary
unzip ("w4_project_data.zip")
```
# Merge training and test data set, with a flag variable for training or test
```{r}
directory<-getwd()

```
# Extracts only the measurements on the mean and std deviation for each measurement 

# Use the descriptive activity names to name the activities in the data set 

# Label the data set with descriptive names

# Create a second independently tidy data set with the average of each variable for each activity and each subject 
