---
title: "w4_project_note_for_submission.Rmd"
output: 
        html_document:
                toc: true
                toc_float: true
                number_sections: true
                theme: darkly
---
## About
This file documents the process working through week 4 course project in data science/getting andcleaning data course. 


## Download the data and store in w4_project_data.zip

```{r download file}

library(downloader)
download("https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip",dest="w4_project_data.zip",mode="wb") #wb stands for write and binary
unzip ("w4_project_data.zip")
```
# Import and consolidate subject,x,y data into sxy_data
## consolidate x_test and x_train with selected features 
### find selected features 
feature_mean_std stores index of selected features
feature_mean_std_names stores index and names of selected features
```{r creates feature_mean_std which stores index of selected features;}
directory<-getwd()
featurefilepath<-paste0(directory,"/UCI HAR Dataset/features.txt")
features<-read.table(featurefilepath,header=FALSE)
colnames(features)<-c("index","featurename")
head(features,4)
#we only need featurenames containing "-mean()-" and "-std()-"
feature_mean_std<-grep("mean|std",features$featurename,ignore.case = TRUE)
length(feature_mean_std)
feature_mean_std_names<-features[feature_mean_std,]
head(feature_mean_std_names)
``` 

### read X_test.txt with selected features 
```{r load X_test.txt and stores in x_test with selected features}
testfilepath<-paste0(directory,"/UCI HAR Dataset/test/")
Xtestfilepath<-paste0(testfilepath,"X_test.txt")
x_test<-read.table(Xtestfilepath,header=FALSE)
dim(x_test)
##next extracting selected features
x_test_backup<-x_test
x_test<-x_test[,feature_mean_std]
dim(x_test)
```
### read X_train.txt with selected features
```{r load X_train.txt and stores in x_train with selected features}
trainfilepath<-paste0(directory,"/UCI HAR Dataset/train/")
Xtrainfilepath<-paste0(trainfilepath,"X_train.txt")
x_train<-read.table(Xtrainfilepath,header=FALSE)
dim(x_train)
x_train_backup<-x_train
x_train<-x_train[,feature_mean_std]
dim(x_train)
```
### union x_test and x_train into x
```{r union test and train with a flag}
library(dplyr)
x_test_backup<-x_test 
x_test<-x_test%>%mutate(source="test")
x_train_backup<-x_train
x_tran<-x_train%>%mutate(source="train")
x<-union_all(x_test,x_train)
str(x)

```
### add feature names to x
```{r add feature names to x}
colnames(x)<-c(as.character(feature_mean_std_names$featurename),"source")
```
**at this point, x_test and x_train consolidation is done**

## consolidate y_test and y_train with labels
### load y_test.txt and store in y_test
```{r load y_test.txt and store in y_test}
Ytestfilepath<-paste0(testfilepath,"y_test.txt")
y_test<-data.frame()
y_test<-read.table(Ytestfilepath,header=FALSE)
head(y_test,3)

```
### load y_train.txt and store in y_train
```{r load y_train.txt and stores in y_train}
ytrainfilepath<-paste0(trainfilepath,"y_train.txt")
y_train<-read.table(ytrainfilepath,header=FALSE)
head(y_train,3)
```
### union y_test and y_train into y
```{r union y_test and y_train}
dim(y_test)
dim(y_train)
y<-data.frame()
#use union_all to maintain all rows
y<-dplyr::union_all(y_test,y_train) 
dim(y)
```

### load activity labels
```{r load activity labels}

## load activity_labels.txt which are the activity associated with 1,2,3,4,5
activityfilepath<-paste0(directory,"/UCI HAR Dataset/activity_labels.txt")
activity_labels<-read.table(activityfilepath,header=FALSE)
colnames(activity_labels)<-c("num","activity")
str(activity_labels)
```
### assign activity names to y
```{r assign activity names to y}
colnames(y)<-c("num")
y_backup<-y #create a backup copy
y<-dplyr::left_join(y,activity_labels,by=c("num"="num"))
head(y)
tail(y)
```
**up to this point, both x and y are clean data

## consolidate subject_test and subject_train into subject 
```{r load subject_test in subject_test}
#load subject_test.txt
Subjecttestfilepath<-paste0(testfilepath,"subject_test.txt")
subject_test<-data.frame()
subject_test<-read.table(Subjecttestfilepath,header=FALSE)
colnames(subject_test)<-c("subject")
```
```{r load subject_train}
Subjecttrainfilepath<-paste0(trainfilepath,"subject_train.txt")
subject_train<-data.frame()
subject_train<-read.table(Subjecttrainfilepath,header=FALSE)
colnames(subject_train)<-c("subject")
head(subject_train)
```
```{r union subject_test and subject_train into subject}
subject<-data.frame()
subject_backup<-subject
subject<-dplyr::union_all(subject_test,subject_train)

```

## consolidate subject,x,y into sxydata
```{r check subject,x and y, paged.print=TRUE}
dim(subject)
dim(x)
dim(y)
colnames(subject)
colnames(y)
colnames(x)
```
```{r consolidate subject,x,y$activity into sxydata}
sxydata<-data.frame()
subject$subject<-as.character(subject$subject)
sxydata<-cbind.data.frame(subject,x,y$activity)
```

# Clean Intertial Signals

List of dataframes generated

* body_acc_x_meansd
* body_acc_y_meansd
* body_acc_z_meansd
* body_gyro_x_meansd
* body_gyro_y_meansd
* body_gyro_z_meansd
* total_acc_x_meansd
* total_acc_y_meansd
* total_acc_z_meansd

## What's in each file? 
what is in each file? 
```{r examine files body_acc_x_test.txt}
body_acc_x_testpath<-paste0(testfilepath,"/Inertial Signals/body_acc_x_test.txt")
body_acc_x_test<-read.table(body_acc_x_testpath,header=FALSE)
#str(body_acc_x_test)
```

```{r examine files body_acc_x_train.txt}
body_acc_x_trainpath<-paste0(trainfilepath,"/Inertial Signals/body_acc_x_train.txt")
body_acc_x_train<-read.table(body_acc_x_trainpath,header=FALSE)
#str(body_acc_x_train)
```

## Create body_acc_x_mean and body_acc_x_sd 
```{r union body_acc_x_test and body_acc_x_train to body_acc_x, in this order}
body_acc_x<-union_all(body_acc_x_test,body_acc_x_train)
dim(body_acc_x)
```
## Extracts only the measurements on the mean and std deviation for each measurement 
```{r compute body_acc_x_mean and body_acc_x_sd}
body_acc_x_backup<-data.frame()
body_acc_x_backup<-body_acc_x
#stores body_acc_x in _backup dataframe
body_acc_x_mean<-rowMeans(body_acc_x,na.rm=TRUE)

#install.packages("matrixStats")
library(matrixStats)
body_acc_x_sd<-matrixStats::rowSds(as.matrix(body_acc_x),na.rm = TRUE)
body_acc_x_meansd<-cbind(body_acc_x_mean,body_acc_x_sd)
colnames(body_acc_x_meansd)<-c("body_acc_x_mean","body_acc_x_sd")
body_acc_x_meansd<-as.data.frame(body_acc_x_meansd)
str(body_acc_x_meansd)
```

## write a function inersig_clean to execute the sequance of:import, union,mean,std calculation

```{r function_inersig_clean}
inersig_clean<-function(iner_test_filename="",iner_train_filename=""){
        #iner_test_filename is the filename of the test file, e.g. body_acc_y_test.txt
        #iner_train_filename is the filename of the train file, e.g.                  body_acc_y_train.txt 
        #the function takes two filenames as the input, load the files, and output a data frame with "mean" and "std" two variables
        # 0. validation
        # 1. Import test and train data         
        iner_testpath<-paste0(testfilepath,"/Inertial Signals/",iner_test_filename)
        iner_testdata<-read.table(iner_testpath,header=FALSE)        
        iner_trainpath<-paste0(trainfilepath,"/Inertial Signals/",iner_train_filename)
        iner_traindata<-read.table(iner_trainpath,header=FALSE) 
        
        # 2. Union two data sets 
        iner_fulldata<-dplyr::union_all(iner_testdata,iner_traindata)
        # 3. Compute mean and std
        iner_mean<-rowMeans(iner_fulldata,na.rm = TRUE) #iner_mean tested ok
        #library(matrixStats)
        iner_sd<-matrixStats::rowSds(as.matrix(iner_fulldata),na.rm = TRUE)#tested ok
        iner_meansd<-cbind(iner_mean,iner_sd)
        colnames(iner_meansd)<-c("mean","sd")
        iner_meansd<-as.data.frame(iner_meansd)
        return(iner_meansd)
        
}

#test for body_acc_x

output<-inersig_clean("body_acc_x_test.txt","body_acc_x_train.txt")
head(output)
head(body_acc_x_meansd)

```
## Use inersig_clean function to combine and clean all inertial signals

List of dataframes generated

* body_acc_x_meansd
* body_acc_y_meansd
* body_acc_z_meansd
* body_gyro_x_meansd
* body_gyro_y_meansd
* body_gyro_z_meansd
* total_acc_x_meansd
* total_acc_y_meansd
* total_acc_z_meansd

each data frame should have two variables, specifying the measure name, e.g "body_acc_x_mean" and "body_acc_y_mean"

### Code
```{r inersig_clean_body_acc_x}
body_acc_x_meansd_backup<-body_acc_x_meansd
body_acc_x_meansd<-inersig_clean("body_acc_x_test.txt","body_acc_x_train.txt")
colnames(body_acc_x_meansd)<-c("body_acc_x_mean","body_acc_x_sd")
str(body_acc_x_meansd)
```

```{r inersig_clean_body_acc_y}
body_acc_ymeansd<-inersig_clean("body_acc_y_test.txt","body_acc_y_train.txt")
colnames(body_acc_ymeansd)<-c("body_acc_y_mean","body_acc_y_sd")
str(body_acc_ymeansd)
```

```{r inersig_clean_body_acc_z}
body_acc_zmeansd<-inersig_clean("body_acc_z_test.txt","body_acc_z_train.txt")
colnames(body_acc_zmeansd)<-c("body_acc_z_mean","body_acc_z_sd")
str(body_acc_zmeansd)
```

```{r inersig_clean_body_gyro_x}
body_gyro_xmeansd<-inersig_clean("body_gyro_x_test.txt","body_gyro_x_train.txt")
colnames(body_gyro_xmeansd)<-c("body_gyro_x_mean","body_gyro_x_sd")
str(body_gyro_xmeansd)
```

```{r inersig_clean_body_gyro_y}
body_gyro_ymeansd<-inersig_clean("body_gyro_y_test.txt","body_gyro_y_train.txt")
colnames(body_gyro_ymeansd)<-c("body_gyro_y_mean","body_gyro_y_sd")
str(body_gyro_ymeansd)
```

```{r inersig_clean_body_gyro_z}
body_gyro_zmeansd<-inersig_clean("body_gyro_z_test.txt","body_gyro_z_train.txt")
colnames(body_gyro_zmeansd)<-c("body_gyro_z_mean","body_gyro_z_sd")
str(body_gyro_zmeansd)
```

```{r inersig_clean_total_acc_x}
total_acc_x_meansd<-inersig_clean("total_acc_x_test.txt","total_acc_x_train.txt")
colnames(total_acc_x_meansd)<-c("total_acc_x_mean","total_acc_x_sd")
str(total_acc_x_meansd)
```

```{r inersig_clean_total_acc_y}
total_acc_y_meansd<-inersig_clean("total_acc_y_test.txt","total_acc_y_train.txt")
colnames(total_acc_y_meansd)<-c("total_acc_y_mean","total_acc_y_sd")
str(total_acc_y_meansd)
```

```{r inersig_clean_total_acc_z}
total_acc_z_meansd<-inersig_clean("total_acc_z_test.txt","total_acc_z_train.txt")
colnames(total_acc_z_meansd)<-c("total_acc_z_mean","total_acc_z_sd")
str(total_acc_z_meansd)
```


# Combine sxydata with the inertial signals into sxy_iner (final data set)

```{r combine_sxy_data_inertial_sig}
sxy_iner<-data.frame()
sxy_iner<-cbind(sxydata,body_acc_x_meansd,body_acc_ymeansd,body_acc_zmeansd,
                body_gyro_xmeansd,body_gyro_ymeansd,body_gyro_zmeansd,
                total_acc_x_meansd,total_acc_y_meansd,total_acc_z_meansd)
colnames(sxy_iner)[89]<-"activity"
write.csv(sxy_iner,file="sxy_iner.csv")
```

# Create a second independently tidy data set with the average of each variable for each activity and each subject 

```{r convert_sxy_iner_to_by_sub}
sxy_iner<-read.csv("sxy_iner.csv",header = TRUE)
sxy_iner<-dplyr::select(sxy_iner,-source,-X)#remove source and index X
sxy_iner$subject<-as.factor(sxy_iner$subject) #turn subject into a factor
sxy_iner_summary<-sxy_iner%>%group_by(subject,activity)%>%summarise_all(mean,na.rm=TRUE)

head(sxy_iner_summary)
write.csv(sxy_iner_summary,file="sxy_iner_summary.csv")

```
