---
title: "w4_quiz"
author: "Tanya Tan"
date created: "27/08/2019"
date updated: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# About
This is an R markdown practice that tracks the codes used in Week 4 Quiz
(*last updated: `r format(Sys.time(), '%d %B, %Y')`*)

# Question 1(Q1)
## Q1.Code
download files
```{r}
download.file("https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv",destfile="housingdata.csv")
housingdata<-read.csv("housingdata.csv",header=TRUE)
```
review data
```{r}
tail(names(housingdata))
```
split names by wgtp and return a list
```{r}
namelist<-strsplit(names(housingdata),"wgtp")
tail(namelist)
namelist[[123]]
```

## Q1.Answer
"" "15"

# Question 2
## Q2.Code
download data to gdp_data and read to csv natively
```{r}
download.file("https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv",destfile="gdp_data.csv")
gdp_data<-read.csv("gdp_data.csv",header=FALSE,sep=",")
head(gdp_data,10)
```
the data is quite messy on the top, and we should start importing from line 6
```{r}
gdp_data<-read.csv("gdp_data.csv",header=FALSE,sep=",",skip = 5)
head(gdp_data,10)
```
check the bottom of the data
```{r}
tail(gdp_data,10)
```
see all NAs at the bottom (remove NA later for mean calculation)
next, select non-NA columns
```{r}
gdp_data<-gdp_data[,c(1,2,4,5)]
head(gdp_data)
```
assign column names
```{r}
colnames(gdp_data)<-c("code","rank","country","gdp")
head(gdp_data)
```
remove the "," in gdp and convert to numeric; note gsub replaces all the , with nothing therefore
it removes the commas
```{r}
gdp_data$gdp<-as.numeric(gsub(",","",gdp_data$gdp))
head(gdp_data)
```

convert rank to numeric (so the blank cells should be NA)
also, if the factor is number you first convert it to a character vector and then to numeric.
```{r}
gdp_data_backup<-gdp_data
class(gdp_data$rank) # this is a factor
gdp_data$rank<-as.numeric(as.character(gdp_data$rank))
head(gdp_data$rank)# this should return 1,2,3,4....
summary(gdp_data$rank)
```


next, only include complete cases
```{r}
dim(gdp_data)
gdp_data<-gdp_data[complete.cases(gdp_data),]
dim(gdp_data)
summary(gdp_data)
```
note, the code and country are still factors; can convert them but for now, keep it is for Q2

compute the mean, removing NA
```{r}
mean(gdp_data$gdp,na.rm = TRUE)
```
## Q2.Answer
answer is `r mean(gdp_data$gdp,na.rm = TRUE)`
# Question 3
## Q3.Code
gdp_data should now contains the information for 190 countries. Create a vector countryNames that stores all country names
```{r}
head(gdp_data)
countryNames<-as.character(gdp_data$country) 
#if not using as.character, it'll be loaded as factor
head(countryNames)
countryNames[grep("^United",countryNames)]
```

## Q3.Answer
the answer is `r length(grep("^United",countryNames)) `

# Question 4
## Q4.Code
the gdp data is already stored in gdp_data;
now download and read the eductional data
```{r}

download.file("https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv",destfile = "edu_data.csv")
edu_data<-read.csv("edu_data.csv",header = TRUE)
dim(edu_data)
head(edu_data)
```
the head seems to be okay. Now tail
```{r}
tail(edu_data)
```
The country code is stored in CountryCode and the "fiscal year ends" wording appears in special notes
```{r}
library(dplyr)
edu_data_short<-select(edu_data,CountryCode,Special.Notes)
head(edu_data_short)
```
search "fical year end: June" and ignore cases
```{r}
edu_data_short[grep("fiscal year end: June",edu_data_short$Special.Notes,ignore.case = TRUE),]
num<-length(grep("fiscal year end: June",edu_data_short$Special.Notes,ignore.case = TRUE))
```
## Q4.Answer
`r num`
# Question 5
load data and save the dataframe in date
```{r load data for q5}
library(quantmod)
library(dplyr)
library(lubridate)
amzn = getSymbols("AMZN",auto.assign=FALSE)
sampleTimes = index(amzn)
date<-ymd(sampleTimes)
date<-as.data.frame(date)
colnames(date)<-"YMD"
head(date)
```
create sampleyear
```{r}
date_backup<-date
date<-date%>% mutate(sampleyear=year(YMD))
head(date)
```
create sampleweekday
```{r}
date_backup2<-date
date<-date %>% mutate(sampleweekday=weekdays(YMD))
head(date)
```
find number of rows in 2012; find number of rows on Mondays
```{r}
record2012<-filter(date,sampleyear==2012)
print("number of records in 2012:")
nrow(record2012)
print("number of records on Monday:")
record2012monday<-filter(date,sampleweekday=="Monday"& sampleyear==2012)# note using & not && here
nrow(record2012monday)
```

notice the difference between & and &&: (p.s. need an empty line before the list * to make the formatting work)

* & and && indicate logical AND and | and || indicate logical OR. The shorter form performs elementwise comparisons in much the same way as arithmetic operators. The longer form evaluates left to right examining only the first element of each vector. Evaluation proceeds only until the result is determined. The longer form is appropriate for programming control-flow and typically preferred in if clauses.
* Long form(&& or ||) short circuits, which means if it can identify the result by just validating just the first element. While doing &&, if the comparision of first two elements resulted in false, comparing next set of elements will also result in False. So, it returns false. While doing || if comparision resulted in true in first few elements, we can confidently say that any further validations will not change the result so it returns True.
* Short forms continues to do for the entire vectors and creates a vector of results and returns it.

## Q5.Answer
year 2012 has `r nrow(record2012)` entries 
year 2012 and monday has `r nrow(record2012monday)` entries

# Notes on Using Rmarkdown
Rmarkdown uses the folder wherer the .Rmd is stored as the working directory; thus, setwd() or getwd()
won't work test here (after forcing the working directory to be elsewhere)
```{r}
getwd()
```