---
title: "w4_quiz"
author: "Tanya tan"
date: "27/08/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# About (# is heading 1)
This is an R markdown practice that tracks the codes used in Week 4 Quiz

# Question 1(Q1)
## Q1.Code
set working directory and download files
```{r}
setwd("~/_GitHub/coursera-data-science-specialization/getting and cleaning data/w4_quiz_data")
download.file("https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv",destfile="housingdata.csv")
download.file("https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FPUMSDataDict06.pdf",destfile = "housingdata_codebook.pdf")
housingdata<-read.csv("housingdata.csv",header=TRUE)
```
review data
```{r}
tail(names(housingdata))
```
split names by wgtp and return a list
```{r}
namelist<-strsplit(names(housingdata),"wgtp")
tail(namelist)
namelist[[123]]
```

## Q1.Answer
"" "15"

# Question 2
## Q2.Code
download data to gdp_data and read to csv natively
```{r}
download.file("https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv",destfile="gdp_data.csv")
gdp_data<-read.csv("gdp_data.csv",header=FALSE,sep=",")
head(gdp_data,10)
```
the data is quite messy on the top, and we should start importing from line 6
```{r}
gdp_data<-read.csv("gdp_data.csv",header=FALSE,sep=",",skip = 5)
head(gdp_data,10)
```
check the bottom of the data
```{r}
tail(gdp_data,10)
```
see all NAs at the bottom (remove NA later for mean calculation)
next, select non-NA columns
```{r}
gdp_data<-gdp_data[,c(1,2,4,5)]
head(gdp_data)
```
assign column names
```{r}
colnames(gdp_data)<-c("code","rank","country","gdp")
head(gdp_data)
```
remove the "," in gdp and convert to numeric; note gsub replaces all the , with nothing therefore
it removes the commas
```{r}
gdp_data$gdp<-as.numeric(gsub(",","",gdp_data$gdp))
head(gdp_data)
```

convert rank to numeric (so the blank cells should be NA)
also, if the factor is number you first convert it to a character vector and then to numeric.
```{r}
gdp_data_backup<-gdp_data
class(gdp_data$rank) # this is a factor
gdp_data$rank<-as.numeric(as.character(gdp_data$rank))
head(gdp_data$rank)# this should return 1,2,3,4....
summary(gdp_data$rank)
```


next, only include complete cases
```{r}
dim(gdp_data)
gdp_data<-gdp_data[complete.cases(gdp_data),]
dim(gdp_data)
summary(gdp_data)
```
note, the code and country are still factors; can convert them but for now, keep it is for Q2

compute the mean, removing NA
```{r}
mean(gdp_data$gdp,na.rm = TRUE)
```
## Q2.Answer
answer is `r mean(gdp_data$gdp,na.rm = TRUE)`
# Question 3
## Q3.Code
gdp_data should now contains the information for 190 countries. Create a vector countryNames that stores all country names
```{r}
head(gdp_data)
countryNames<-as.character(gdp_data$country) 
#if not using as.character, it'll be loaded as factor
head(countryNames)
countryNames[grep("^United",countryNames)]
```

## Q3.Answer
the answer is `r length(grep("^United",countryNames)) `

# Question 4
## Q4.Code
## Q4.Answer