---
title: "Simulation Report"
author: "Tanya T"
date: "10/10/2019"
output:
  pdf_document: default
  html_document: default
---
# Overview

This report is divided in part 1 and part 2. 

Central Limit Theorem (CLT) states that the sampling distribution of t he sample means approaches a normal distribution as the sample size gets larger. The goal of part 1 is to compare the simulated exponential distribution in R with the theoretical distribution. With a large number 1000, we are going to prove the the simulated sample produces a similar distribution with the theoretical distribution 

In part 2, we examined the ToothGrowth data, and compared the means of the two supplment delivery types (ascorbic acid, coded as VC and orange juice, coded as OJ); as well as the tooth growths among three difference doses (0.5,1 and 2 mg/day). We conclude that, there is no difference in toothgrowth between two delivery methods, by orange juice or ascorbic acid. However, the higher the dose is, the more tooth growth is seen. 

## Load Libraries
```{r}
library(ggplot2)

```


# Part 1 Simulation

Here we invesitgatethe exponential distribution in R and compare it with the Central Limit Theorem (CLT. 

## More on Exponential Distribution

The exponential distribution is concerned with the amount of time until a specific event occurs. e.g. the amount of time until the next rain storm likely has an exponential distribution. R's exp function takes two arguments: 

        * The number of observations you want to see 
        * The estimated rate of events for the distribution: this is usually 1/expected service life or wait time

## Simulation

First, we set the initial values for lambda, sample size, and number of simulations

```{r setup}
set.seed(1) # set seed for reprocibability
lambda<-0.2
nsample<-40
simn<-1000 # number of simulations
```
The simualted data is a 40 by 1000 matrix. 

```{r simdata}
simdata<-replicate(simn,rexp(nsample,lambda)) #create simdata matrix 
str(simdata)
```

## Sample Mean versus Theoretical Mean 

Next, the sample mean is calculated
```{r}
#sim_mean is the mean of simulated data
sim_mean<-colMeans(simdata) #calculate column means, each column is 40 samples 
str(sim_mean)

#The theoretical mean and standard deviation are calculated 
#theoretical mean and SD
theo_mean<-1/lambda
theo_sd<-theo_mean/sqrt(nsample)
theo_var<-theo_sd^2
```
The mean of simulation is `r mean(sim_mean)` whereas the theoretical mean is `r theo_mean`, as shown in the graph below

```{r compmean}
sim_mean<-as.data.frame(sim_mean)
colnames(sim_mean)<-"value"
ggplot(data = sim_mean,aes(x=value))+geom_histogram(binwidth = 0.2)+
        geom_vline(xintercept = theo_mean,colour=c("red"),linetype="dashed")+labs(title="Comapring Sample Mean Distributions and the Theoretical Mean",x="Mean Values",y="Frequency")

```


## Sample Variance versus Theoretical Variance
```{r}
sim_var<-sd(as.matrix(sim_mean))^2
sim_var
```

The theoretical variance is `r theo_var` whereas the sample variance is `r sim_var`


## Distribution

Next we will show the distribution is approximately normal by overlaying a normal distribution with the theoretical mean and standard deviation
```{r}
#note here we need density function for geom_histogram

ggplot(data = sim_mean,aes(x=value))+
        geom_histogram(aes(y=..density..),binwidth = 0.2)+ 
        stat_function(fun=dnorm,args=list(mean=theo_mean,sd=theo_sd),color="green",
                      lwd=3,linetype="dashed")+labs(title = "Comparing the Sample Distribution with the Normal Density Curve",xlab="Mean Values")


```

Based on the plot above that we know the distribution is approximately normal

# Part 2 Basic Inferential Data Analysis

## Load ToothGrowth Data
```{r}
data("ToothGrowth")
str(ToothGrowth)
```

We examined the ToothGrowth data and found it contains three variables, len, supp and dos. 
        *len: the numeric tooth length 
        *supp: supplement type (VC or OJ)
        *dose: numeric dose in milligrams/day 


Next we visualize the data 
```{r}
ggplot(data = ToothGrowth,aes(x=supp,y=len))+
        geom_boxplot()+labs(title = "Tooth Lengths by Delivery Method of Vitamin C",
                            x="Delivery Method",y="Tooth Length")
```


```{r}
ggplot(data = ToothGrowth,aes(x=dose,y=len,color=supp))+
        geom_boxplot()+labs(title = "Tooth length by Vitamin Doses",x="Dose (mg/day)",
                            y="Tooth Lengh")
```

## Compare tooth growth by supp
```{r}
t.test(len~supp,data = ToothGrowth)
```
since the p value is greater than 0.05, there is no statistically significant difference on toothgrowth by supplement types

## Compare tooth growth by dose 
```{r}
ToothGrowth$dose<-as.character(ToothGrowth$dose)
res<-aov(len~dose,data = ToothGrowth)
summary(res)
```
A one-way ANOVA test was performed to examine the differences of tooth lenghs by dose groups. The p value indicates that there is a statistically detectable difference. 

In addition, Tukey pairwise comparisons are conducted to examine the differences between each pair of groups 

```{r}
TukeyHSD(res)
```
From the results, we've learned that the differences are all significant. 

## Conclusion

        * The assumptions for conducting these two-sample t-tests are that the groups are all independent of each other. 
        *Although the mean of OJ group(`r mean(ToothGrowth$len[ToothGrowth$supp=="OJ"])`) is greater than the mean tooth length of VC group (`r mean(ToothGrowth$len[ToothGrowth$supp=="VC"])`), such difference is not statistically significant; However, the 
        *There are, however, statistically detectable differences among the groups with different doses. Thet mean values and confidence intervals for each dose group are: 
                - Dose=0.5 `r mean(ToothGrowth$len[ToothGrowth$dose=="0.5"])` with the confidence interval `r t.test(ToothGrowth$len[ToothGrowth$dose=="0.5"])$conf.int`
               
                - Dose=1 `r mean(ToothGrowth$len[ToothGrowth$dose=="1"])` with the confidence interval `r t.test(ToothGrowth$len[ToothGrowth$dose=="1"])$conf.int`
                - Dose=2 `r mean(ToothGrowth$len[ToothGrowth$dose=="2"])` with the confidence interval `r t.test(ToothGrowth$len[ToothGrowth$dose=="2"])$conf.int`

The higher the dose is, the more tooth growth there is. 



